Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing

### 2019-10-19 Dissonance_Santiago_v1.Rmd
Edited the abstract and updated the README.md file

---
title: Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Increasingly, however, it is recognized that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain users of public transport and active travel. A challenge with the way transportation policy is developed is that it tends to focus on cost-benefit and utilitarian considerations, which may miss other potential benefits of transportation. For this reason, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective and utilitarian values. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective and utilitarian values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associated to affective values (freedom, enjoyment, and happiness) and to utilitarian values (safety, time waste, and efficiency). The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with affective values.

  To properly assess transportation policy This poses a number of challenges Broadening the commonly used transport project assessment will be relevant for achieving this goal. While satisfaction with travel has been mainly used in the context of daily trips and mainly under cost-benefit and utilitarian considerations, subjective wellbeing (SWB) over time has risen as an alternative for understanding the preferences and choices of transport users. The aim of this paper is to analyse the factors that can potentially affect the levels of SWB in the wide spectrum of public, private and active transport modes in the context of Santiago de Chile. The paper aims to reflect on the possibilities of increasing the SWB with a view to enhance the attractiveness of public transport and active travel. The paper investigates the affective and instrumental/utilitarian factors of SWB, according to modal choice. It uses an analysis of correspondence that considers the actual primary mode used, versus the mode that each individual associated to both utilitarian and affective SWB factors. The results show that there is more persistent modal dissonance between positive emotional and utilitarian associations within the more disadvantaged groups – particularly among users of public transport. The study shows that apart from time-savings and efficiency, security has also been cross-sectionally sensitive according to the data analysed. Factors such as happiness, enjoyment and freedom require far more attention in transport planning as they are strongly related to the dimensions of subjective and personal wellbeing. Targeting these factors will be relevant for attracting new users to public transportation and active travel.

@@

# To run this notebook, the file `Dissonance_Santiago.RData` is needed. This file contains the following object:
# 
# master: a dataframe with information collected from 451 respondents. The variables in the dataframe are:
# 
# - NUMERO: Respondent ID
# - USED: Main mode of transportation used by the respondent
# - FREEDOM: Mode(s) that the respondent associates with feelings of freedom
# - ENJOYMENT: Mode(s) that the respondent associates with feelings of enjoyment
# - HAPPINESS: Mode(s) that the respondent associates with feelings of happiness
# - POVERTY: Mode(s) that the respondent associates with feelings of poverty
# - LUXURY: Mode(s) that the respondent associates with feelings of luxury
# - STATUS: Mode(s) that the respondent associates with feelings of status
# - SAFETY: Mode(s) that the respondent associates with feelings of safety
# - TIME: Mode(s) that the respondent associates with feelings of WASTED TIME
# - EFFICIENCY: Mode(s) that the respondent associates with feelings of efficiency
# - AGE: Age of the respondent
# - EDUCATION: Level of education of the respondent
# - INCOME: Level of income of respondent
# - TRAVEL_TIME: Trip duration of regular commuting trip 
# - WEIGHT_*: A weight for the responses by ATTRIBUTE. Some respondents selected more than one mode that they associated with feelings of {FREEDOM, ENJOYMENT, HAPPINESS, SAFETY, TIME, and EFFICIENCY}. To avoid giving greater weight to respondents who indicated several modes of transportation, a weight was calculated that is the inverse of the number of modes indicated. These weights can be used to weight the frequencies prior to any analysis.
# 
# The modes of transportation (USED, FREEDOM, ENJOYMENT, HAPPINESS, SAFETY, TIME, EFFICIENCY) are coded as: 
# "A": Auto
# "B": Taxi
# "C": Colectivo
# "D": Moto
# "E": Metro
# "F": Bus
# "G": Bicycle
# "H": Walk
#
# AGE is coded as:
#
# "A": Less18
# "B": 18to24
# "C": 25to34
# "D": 35to54
# "E": 55to64
# "F": 65more
# 
# EDUCATION is coded as:
# 
# "A": ELEMENTARY
# "B": SECUNDARY
# "C": PROFESS_TECH
# "D": COLLEGE
# "E": POSTGRAD
#
# INCOME is coded as (in pesos?):
# 
# "A": Less423
# "B": 423to639
# "C": 639to977
# "D": 977to1550
# "E": 1550to2380
# "F": 2380more
# 
# TRAVEL_TIME is coded as:
#
# "A": 0-20 min
# "B": 20-40 min
# "C": 40-60 min
# "D": 1h and more

>>>>

---
title: Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Increasingly, however, it is recognized that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain active travelers and users of public transportation. A challenge with the way transportation policy is developed is that it tends to focus on utilitarian considerations, which may miss other potential aspects of transportation that users value. In particular, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective values associated with SWB. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective and utilitarian values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associated to affective values freedom, enjoyment, happiness, poverty, luxury and status. The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, public transportation is most commonly associated with poverty, and the automobil is most commonly associated with luxury and status.

@@

# To run this notebook, the file `Dissonance_Santiago.RData` is needed. This file contains the following object:
# 
# master: a dataframe with information collected from 451 respondents. The variables in the dataframe are:
# 
# - NUMERO: Respondent ID
# - USED: Main mode of transportation used by the respondent
# - FREEDOM: Mode(s) that the respondent associates with feelings of freedom
# - ENJOYMENT: Mode(s) that the respondent associates with feelings of enjoyment
# - HAPPINESS: Mode(s) that the respondent associates with feelings of happiness
# - POVERTY: Mode(s) that the respondent associates with feelings of poverty
# - LUXURY: Mode(s) that the respondent associates with feelings of luxury
# - STATUS: Mode(s) that the respondent associates with feelings of status
# - SAFETY: Mode(s) that the respondent associates with feelings of safety
# - TIME: Mode(s) that the respondent associates with feelings of WASTED TIME
# - EFFICIENCY: Mode(s) that the respondent associates with feelings of efficiency
# - AGE: Age of the respondent
# - EDUCATION: Level of education of the respondent
# - INCOME: Level of income of respondent
# - TRAVEL_TIME: Trip duration of regular commuting trip 
# - WEIGHT_*: A weight for the responses by ATTRIBUTE. Some respondents selected more than one mode that they associated with feelings of {FREEDOM, ENJOYMENT, HAPPINESS, POVERTY, LUXURY, STATUS, SAFETY, TIME, and EFFICIENCY}. To avoid giving greater weight to respondents who indicated several modes of transportation, a weight was calculated that is the inverse of the number of modes indicated. These weights can be used to weight the frequencies prior to any analysis.
# 
# The modes of transportation (USED, FREEDOM, ENJOYMENT, HAPPINESS, POVERTY, LUXURY, STATUS, SAFETY, TIME, EFFICIENCY) are coded as: 
# "A": Auto
# "B": Taxi
# "C": Colectivo
# "D": Moto
# "E": Metro
# "F": Bus
# "G": Bicycle
# "H": Walk
#
# AGE is coded as:
#
# "A": Less18
# "B": 18to24
# "C": 25to34
# "D": 35to54
# "E": 55to64
# "F": 65more
# 
# EDUCATION is coded as:
# 
# "A": ELEMENTARY
# "B": SECUNDARY
# "C": PROFESS_TECH
# "D": COLLEGE
# "E": POSTGRAD
#
# INCOME is coded as (in pesos?):
# 
# "A": Less423
# "B": 423to639
# "C": 639to977
# "D": 977to1550
# "E": 1550to2380
# "F": 2380more
# 
# TRAVEL_TIME is coded as:
#
# "A": 0-20 min
# "B": 20-40 min
# "C": 40-60 min
# "D": 1h and more
